### 2013-09-13 report/report.tex
Modified list of papers, report and gitignore

\documentclass[11pt,a4paper]{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{todonotes}

\usepackage{amsmath,amssymb,amsfonts}

\title{Local Adaptive Optimization of Timestep}
\author{Malte Stær Nissen\\University of Copenhagen\\\texttt{malte.nissen@gmail.com} \and Supervisor: Sune Darkner\\University of Copenhagen\\\texttt{darkner@diku.dk}}

\begin{document}
\maketitle

\tableofcontents

\clearpage

\section{Introduction}
Simulations of large (order of millions of vertices) hyperelastic materials are
very time consuming. When working with materials of varying density of vertices,
the highest densities often make the most strict bounds on size of the timestep
of the simulation in order to keep the materials (and grids) stable. In a
``standard'' simulation with a global timestep size we need to compute possibly
unneccessarily many computations on the coarser parts of the material caused by this bound. When
furthermore having materials with a large amount of somewhat equally distributed
vertices and smaller patches of detailed areas with higher density of vertices,
it would be an advantage to be able to perform large timesteps for the majority of
the material and smaller steps for the local patches when vertices cause
instability. We call this concept local adaptive optimization of timestep, which
we will study further in this report.

\section{Previous work}
As according to \cite{Gander:2013} local time stepping was first studiet
in the community of ordinary differential equations (ODE) with Rice
\cite{rice:1960} developing the split Runge-Kutta methods (multirate
Runge-Kutta methods). The concept behind these methods is the splitting of
the ODEs into multiple (two components in \cite{rice:1960}) components of
which each component needs to be integrated in different scales. A strategy
is chosen as proposed in \cite{Kvaernoe:1999} of either computing the coarser
part first (``Fastest first strategy'') followed by the finer part or vice
versa (``Slowest first strategy''). In order to perform these sequential
computations either interpolation or extrapolation of the first computation
is performed to compute the second part in each time step and the different
time step sizes can be adapted to fit the model in each step as well. See
\cite{Kvaernoe:1999} and \cite{Gear:1984} (similar strategy for linear
multistep methods) for more detailed descriptions of the two strategies.

In the partial differential equations (PDE) community, the adaptive
timestepping area was explored and developed with the goal of simulating
very specific known problems such as the (hyperbolic) wave equations
and (parabolic) heat equations instead of more general applications. We
will adhere from describing the methods further in this report, but a
short description and comparison of the specific methods can be found in
\cite{Gander:2013}. Generally the methods developed at first all make use
of the same basic concepts for making the local adaptive timestepping:
Interpolation, extrapolation, prediction and correction, which we are going to
use in the work of this report as well.

Recently newer and faster methods with local time stepping have been
developed on the basis of nonlinear PDEs and the work performed in this
field in the 80s and 90s. First the multiresolution (MR) schemes for creating
space-adaptive discretizations and refinements of these were developed,
see \cite{Berger:1984}. Then multiple MR based schemes were developed.
Dominques et al. \cite{Domingues:2008} is one of the more recent of these MR
schemes. They describe a local scale-dependent time stepping
for a space-adaptive multiresolution scheme using the finite volume method
in order to obtain speed-up using larger time-steps without violating a
defined stability constraints, which is essentially the same motivation as
this report. The method is based on an explicit Runge-Kutta method of second
order. As expected the timestep size is imposed by a stability condition of
the explicit Runge-Kutta on the finest scale, which increases with the scale
of the mesh and hence we are able to increase the timestep as well without
violating the stability condition.

\section{Method}

\section{Experiments and results}

\section{Discussion}

\section{Conclusion}

\cite{Keshav:2007}
\bibliography{bibliography}
\bibliographystyle{plain}

\end{document}