### 2013-10-22 review/review.tex
Initiated review

\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{amsmath,amssymb,amsfonts}

\title{Review of Human Pose Tracking}
\author{Malte Stær Nissen}

\begin{document}
\maketitle

\noindent Corrections/Questions:
\begin{itemize}
    \item 1.L2: Comma before ``and video games'' (Oxford comma)
    \item 1.L13 and 2.4.1.L2: Gradient ascent/descent - which is it? Looks as if you refer to
        the same paper and the same part of the algorithm.
    \item Figure 1: Is the figure your own product or from somewhere else?
    \item 2.1.L3, 2.1.L11: Shouldn't you always start with a capital letter after colon?
    \item 2.3.2.L2: ``requires a shape model, described above'' (where exactly?)
    \item 2.3.3: Any reference? When you have chopped the section into
        subsections it might be a good idea to have references for all
        methods
    \item 2.6.L3: ``the then state-of-the-art method''. The ``then''-part Doesn't sound correct
        in my head/mind. I have however no alternative solution at hand.
\end{itemize}

\noindent General:
\begin{itemize}
    \item Introduction: Good and sufficient describtion to get a grasp of the
        problem and the work being performed (exclusing results).
    \item When reading the litterature study (all the subsections following the
        introduction) it first seems quite messy with a
    lot of terminology but no red line. After a few times of reading the
    subsections and introduction (2.1)
    it gets more obvious what your intention/red line of the litterature study
    is. Maybe emphasize in 2.1.L7 that the overview is the other subsections of
    section 2. Would've helped me a bit when first reading the report.

    \item Quite long litterature study explaining a lot of the therminology
        within the field. Maybe some of it is more suitable in the method
        section?
    \item Consistency in writing style (sections): Some places (2.4) you describe (name) the
        following subsections in the overall section. Other places (2.3.2, 2.3,
        2.5) you only write that there are X (number) of methods/subsections,
        not what they are.
    \item Maybe some sort of conclusion of the litterature study could be added
        at the end? Some comments on what you're going to implement/use of all
        the stuff you present.
\end{itemize}

\end{document}